PANCHANGE_data_preprocessing

### 2020-08-19 PANCHANGE_analysis.Rmd
convert to long format

>>>>

```{r vector with psychiatric disorders}
disorder_columns <- c(
    "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "psychotic_disorders_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric"
)
```

```{r gather to long format}
dat.long <- dat %>%
  gather(
    key = "Disorder",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Disorder,
    Diagnosis
  )

head(dat.long)
```

Define disorder columns in vector
```{r}
dat.long %>%
  group_by(Disorder) %>%
  count()
```

## Disorders - Frequencies *use this as example to convert chunk above*
```{r Disorders - Frequencies}
#0 = No to disorder; 1 = Yes to disorder
PMAD_na_drop_Dx %>%
  select(ID, #need to add columns for which you want summaries for - ASK GB & KY what they want in descriptives table
         dem.sex,
         starts_with("mhd.") & ends_with("numeric")) %>%
  gather(
    key = "Disorder",
    value = "Diagnosis",
    mhd.MDD_numeric:mhd.ADHD_numeric #all the columns that we want to be gathered
  ) %>%
  group_by(Disorder) %>%
  count(Diagnosis) #count for diagnoses
```

Compare the GLAD/EDGI and COPING cohorts as a whole
+++CH: Diagnosis, ethnicity, gender, 
+++MD: May not be possible for NBR; potentially send them code to run it for us

Ethnicity
```{r}

```

Gender
```{r}

```

Education
```{r}

```

Key worker status
```{r}

```

COVID test positive
```{r}

```

COVID probable diagnosis
```{r}

```

Calculate mean change per disorder group 
Disorders (9) 
Sex 
Age (13; RAMP; potentially collapse above 65 depending on group sizes) 
Ethnicity groups (5; White European, Black (British), Asian, Arab, Mixed)

Rank the disorders by the size of the change

# Regression
Regress covariates out of the change? Which?

Model 0: Disorders
```{r}

```

Model 1: Model 0 + Sex + Age + Ethnicity + Time registration and baseline
```{r}

```

Re-calculate change
```{r}

```

Re-rank the change again
```{r}

```

PCL:  
Calculate the total GLAD, COPING, RAMP baseline 
```{r}

```

For each item 
present percentages of how say that this is due to the pandemic 
```{r}

```

Calculate a variable PCL pandemic score 
```{r}

```

If one item is highly endorsed due to pandemic -> exclude in sensitivity analysis 
```{r}

```

Descriptives for reporting: "How different are these feelings to how you felt before the pandemic?’ on a 5-point scale (much worse, a little worse, no different, a little better, much better)."
+++KLP: Discuss during the meeting if we use 3 or 5 categories
```{r}

```

Replicate chunk above for retrospective reports. This will only be for people who say that have reported that there is a change in their scores. 
*Discuss if we should include people who reported "no change" as repition of their current GAD, PHQ, PCL report*
```{r}

```

```{r}

```

### 2020-08-19 PANCHANGE_data_preprocessing2020-07-15.Rmd
convert to long format

### Select variables
```{r edgi employment cleaning}
if(EDGI == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

edgi.employment.raw.id <- edgi.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## RAMP - NOT YET HAVE THE RAMP PCL DATA READY
```{r coping ramp pcl cols}
if(RAMP == TRUE) {
coping.ramp.pcl.raw <- readRDS(file = paste0(data.raw_path,"/ramp/pcl_ramp.rds"))
dim(coping.ramp.pcl.raw)
colnames(coping.ramp.pcl.raw)
}
```

@@

*Note that the COPING GLAD OCI-R ID numbers are stored under recipientFirstName (rather than ExternalDataReference)
```{r coping glad ocir cleaning}
if(GLAD == TRUE) {

@@

*Note that the COPING EDGI OCI-R ID numbers are stored under recipientFirstName (rather than ExternalDataReference)
```{r coping edgi ocir cleaning}

@@

  separate(externalDataReference, into = c("Sample", "ID"), sep = 6) %>% # Split ID in Sample and Number

@@

@@

@@

glad/edgi data frame list
```{r glad/edgi data frame list}
if(GLAD == TRUE | EDGI == TRUE) {
dfs.list <- list(
  dem.raw.id,
  employment.raw.id,
  mhd.raw.categories,
  prepan.gad.raw.id,
  prepan.pcl.raw.id,
  prepan.phq.raw.id,
  prepan.ocir.raw.id,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
}
```

@@

Results:
Persistent cough: "Mild symptoms" as cut off
Fatigue: "Severe symptoms" as cut off
Skipped meals: "Moderate symptoms" as cut off
```{r proportion of people who report COVID symptoms in Menni et al}
#persistent cough
percentage_pcough <- ((0.5673*6452) + (0.4556*9186) + (0.3017*2434931))/(6452+9186+2434931)
percentage_pcough    # 0.3029762% of Menni et al(2020) have fatigue
quantile(x = respiratory.raw.id$respiratory.persistent_cough_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_pcough) # this gives 0.3029762% of the sample reporting fatigue when using the "mild symptoms" persistent cough response

#fatigue
percentage_fatigue <- ((0.298*6452) + (0.1556*9186) + (0.0871*2434931))/(6452+9186+2434931) 
percentage_fatigue   # 0.08791204% of Menni et al(2020) have fatigue
quantile(x = respiratory.raw.id$respiratory.feeling_tiredfatigued_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_fatigue) #this gives 8% of the sample reporting fatigue when using the "severe symptoms" fatigue response

#skipped meals 
percentage_skipped_meals <- ((0.4203*6452) + (0.2493*9186) + (0.1912*2434931))/(6452+9186+2434931) 
percentage_skipped_meals    # 0.192021% of Menni et al(2020) have fatigue
quantile(x = respiratory.raw.id$respiratory.skipping_meals_because_you_felt_unwell_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_skipped_meals) # this gives 0.192021% of the sample reporting fatigue when using the "moderate symptoms" persistent cough response
```

@@

#persistent cough
respiratory.raw.id <- respiratory.raw.id %>%
  mutate(
    respiratory.pcough_binary =
      if_else(
        respiratory.persistent_cough_numeric >= 1, #those who score 1 or above (mild or above)
        1,
        0
      )
  )

#fatigue
respiratory.raw.id <- respiratory.raw.id %>%
  mutate(
    respiratory.fatigue_binary =
      if_else(
        respiratory.feeling_tiredfatigued_numeric >= 3, #those who score 3 (severe symptoms)
        1,
        0
      )
  )

#skipped meals
respiratory.raw.id <- respiratory.raw.id %>%
  mutate(
    respiratory.skipped_meals_binary =
      if_else(
        respiratory.skipping_meals_because_you_felt_unwell_numeric >= 2,
        1,
        0
      )
  )
```

>>>>

### Select variables
```{r edgi employment cleaning}
if(EDGI == TRUE){
exclude_cols_employment_edgi <- c("ID",
                  "Sample",
                  "COPING_ID")

edgi.employment.raw.id <- edgi.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  separate("Login ID", into = c("COPING_Sample", "COPING_ID"), sep = 6) %>% # Split ID in Sample and Number
  mutate(COPING_ID = as.numeric(COPING_ID)) %>%
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         COPING_ID,
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment_edgi) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## RAMP
```{r coping ramp pcl cols}
if(RAMP == TRUE) {
coping.ramp.pcl.raw <- readRDS(file = paste0(data.raw_path,"/ramp/pcl_ramp.rds"))
dim(coping.ramp.pcl.raw)
colnames(coping.ramp.pcl.raw)
}
```

@@

*Note that the COPING GLAD OCI-R ID numbers are stored under recipientFirstName (rather than ExternalDataReference)
```{r coping glad ocir cleaning}
if(GLAD == TRUE) {

@@

*Note that the COPING EDGI OCI-R ID numbers are stored under recipientFirstName (rather than ExternalDataReference)
```{r coping edgi ocir cleaning}
exclude_cols_coping_edgi_ocir <- c("COPING_Sample", "COPING_ID")

@@

  separate(externalDataReference, into = c("COPING_Sample", "COPING_ID"), sep = 6) %>% # Split ID in Sample and Number
  mutate(COPING_ID = as.numeric(COPING_ID)) %>%

@@

#    Sample, # Sample
    COPING_ID, # ID

@@

  add_column(Sample = "EDGI") %>%
  add_numeric(., exclude = exclude_cols_coping_edgi_ocir) %>%

@@

```{r edgi employment and coping ocir}
if(EDGI == TRUE){
edgi.employment.coping.ocir <-  full_join(
    x = employment.raw.id,
    y = coping.ocir.raw.id,
    by = 
      c("Sample",
        "COPING_ID")
)
}
```

edgi data frame list
```{r edgi data frame list}
if(EDGI == TRUE) {
dfs.list <- list(
  dem.raw.id,
  edgi.employment.coping.ocir,
  mhd.raw.categories,
  prepan.gad.raw.id,
  prepan.pcl.raw.id,
  prepan.phq.raw.id,
  prepan.ocir.raw.id,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id
#  coping.ocir.raw.id
)
}
```

glad data frame list
```{r glad/edgi data frame list}
if(GLAD == TRUE) {
dfs.list <- list(
  dem.raw.id,
  employment.raw.id,
  mhd.raw.categories,
  prepan.gad.raw.id,
  prepan.pcl.raw.id,
  prepan.phq.raw.id,
  prepan.ocir.raw.id,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
}
```

@@

Results:
Persistent cough: "Mild symptoms" as cut off
Fatigue: "Severe symptoms" as cut off
Skipped meals: "Moderate symptoms" as cut off
```{r proportion of people who report COVID symptoms in Menni et al}
#persistent cough
percentage_pcough <- ((0.5673*6452) + (0.4556*9186) + (0.3017*2434931))/(6452+9186+2434931)
percentage_pcough    # 0.3029762% of Menni et al(2020) have fatigue
pcough_cutoff <- quantile(x = respiratory.raw.id$respiratory.persistent_cough_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_pcough) # this gives 0.3029762% of the sample reporting fatigue when using the "mild symptoms" persistent cough response
pcough_cutoff

#fatigue
percentage_fatigue <- ((0.298*6452) + (0.1556*9186) + (0.0871*2434931))/(6452+9186+2434931) 
percentage_fatigue   # 0.08791204% of Menni et al(2020) have fatigue
fatigue_cutoff <- quantile(x = respiratory.raw.id$respiratory.feeling_tiredfatigued_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_fatigue) #this gives 8% of the sample reporting fatigue when using the "severe symptoms" fatigue response
fatigue_cutoff

#skipped meals 
percentage_skipped_meals <- ((0.4203*6452) + (0.2493*9186) + (0.1912*2434931))/(6452+9186+2434931) 
percentage_skipped_meals    # 0.192021% of Menni et al(2020) have fatigue
skipped_meals_cutoff <- quantile(x = respiratory.raw.id$respiratory.skipping_meals_because_you_felt_unwell_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_skipped_meals) # this gives 0.192021% of the sample reporting fatigue when using the "moderate symptoms" persistent cough response
skipped_meals_cutoff
```

@@

#persistent cough
respiratory.raw.id <- respiratory.raw.id %>%
  mutate(
    respiratory.pcough_binary =
      if_else(
        respiratory.persistent_cough_numeric >= pcough_cutoff, #those who score 1 or above (mild or above)
        1,
        0
      )
  )

#fatigue
respiratory.raw.id <- respiratory.raw.id %>%
  mutate(
    respiratory.fatigue_binary =
      if_else(
        respiratory.feeling_tiredfatigued_numeric >= fatigue_cutoff, #those who score 3 (severe symptoms)
        1,
        0
      )
  )

#skipped meals
respiratory.raw.id <- respiratory.raw.id %>%
  mutate(
    respiratory.skipped_meals_binary =
      if_else(
        respiratory.skipping_meals_because_you_felt_unwell_numeric >= skipped_meals_cutoff,
        1,
        0
      )
  )
```
